# Single-Dataset GIS Operations

## Learning Objectives

- Become familiar with several common single-dataset GIS operations
- Calculate centroids of polygons
- Create buffers
- Explore additional single-dataset GIS operations

## Functions Learned
- `st_geometry()`
- `st_centroid()`
- `st_buffer()`
- `st_coordinates()`
- `st_bbox()`

## Interactive Tutorial

```{block type="rmdinfo"}
This workshop's script can be found [here](https://github.com/spatialanalysis/workshop-scripts/blob/master/R/02-gis-1.R).
```

## Exercises

- Project 1986 ward data into correct UTM projection
```{r}
library(sf)
ward86 <- st_read("data/ward1986.shp")
ward86 <- st_transform(ward86, 32616)
plot(ward86)
```

- Calculate centroids of wards with `st_centroid`
```{r}
?st_centroid

centroids <- st_centroid(ward86)
plot(st_geometry(centroids), cex = 0.1)
plot(st_geometry(ward86), add = T)
```

```{r}
plot(st_geometry(ward86))
plot(st_geometry(centroids), cex = 0.1, add = T)
```

- Calculate bounding box with `st_bbox`
- Plot centroids, buffered centroids, and wards for each year

## Links
- Geometric unary operations (aka, single dataset operations): https://r-spatial.github.io/sf/reference/geos_unary.html
- sf cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/sf.pdf
- PostGIS cheatsheet (off of which sf is based): http://www.postgis.us/downloads/postgis21_cheatsheet.pdf