# Spatial Autocorrelation with rgeoda

This section of workshop notes is based off of the [rgeoda documentation](https://rgeoda.github.io/rgeoda-book/index.html). This package is still in alpha, so the functionality may change. However, you are free to use it for a project for Luc's class.

## Install packages

Install the alpha version of `rgeoda`:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("lixun910/rgeoda")
```

This will install a package that has the same internals as GeoDa, but can be accessed from R. It exposes a lot of functionality that can be hard to manage in R.

You should now be able to load the package with:

```{r}
library(rgeoda)
```

## R Markdown

We'll be working in a R Markdown document today instead of a script. This allows for "literate code" (aka, you can write text along with code - it's how these workshop notes are written). Learn more about R Markdown with [this tutorial](https://swcarpentry.github.io/r-novice-gapminder/15-knitr-markdown/).

To create a new R Markdown file, click **File > New File > R Markdown**. 

The R Markdown cheatsheet and guide can be found under **Help > Cheatsheets > R Markdown Cheat Sheet / Reference Guide**.

## Load data

We'll first load in sample data from the rgeoda package.

```{r}
library(geodaData)
library(rgeoda) # if this doesn't work, have you installed rgeoda from Github?

guerry_sf <- geodaData::guerry
```

```{block type="learncheck"}
**Question**
```
Take a few minutes and try to understand what this data is. 

- How many observations and variables are there? What data is stored? (`dim()`, `str()`, `head()`, `summary()`) 
- What does the metadata tell you about this data? (`?guerry`)
- What geometries are in this data? Can you make a quick map with `plot()`?
- What coordinate reference system is there? Is this data projected? (`st_crs()`) 

```{block type="learncheck"}
```

```{block type="learncheck"}
**Question**
```
Try to make a simple map of one of the attributes of interest with tmap. Can you make it interactive?

```{block type="learncheck"}
```

## Working with rgeoda

```{block type="rmdinfo"}
This workshop is based off of the [rgeoda documentation book](https://rgeoda.github.io/rgeoda-book/). rgeoda is still in alpha, so please let me know if you run into anything funky! You can also record any issues you have on the [Github repository](https://github.com/lixun910/rgeoda/issues).
```

### Convert to rgeoda object

We can convert an sf object into a rgeoda object:

```{r}
guerry <- sf_to_geoda(guerry_sf)
guerry
```

Where did our data go? By default, the sf-to-rgeoda conversion removes the attributes (for faster processing with queen weights, local Moran, etc.). Set `with_table()` to `TRUE` to get the dataframe.

```{r}
guerry <- sf_to_geoda(guerry_sf, with_table = TRUE)
guerry
```

```{block type="rmdinfo"}
Alternatively, you can read a shapefile, geojson (and other spatial file formats) directly into R, with `geoda_open()`:

```{r eval=FALSE}
chicago_libraries <- geoda_open("https://data.cityofchicago.org/resource/x8fc-8rcq.geojson")
chicago_libraries
```
```

### Looking at the rgeoda object

Currently, rgeoda is structured more like Python; we may change this in a future implementation but it's what we have so far. 

Try out the following:

```{r}
guerry$field_names
guerry$field_types
guerry$n_cols
guerry$n_obs
guerry$table
```

Note that this is not really a data frame - rgeoda right now is optimized to help you do spatial weights, autocorrelation and clustering. 

So let's do those things!

## Spatial Analysis with rgeoda

Let's create weights from an rgeoda object:

```{r}
queen_weights(guerry)
```

Much easier than spdep, eh?

You can change the order from first to second:

```{r}
queen_weights(guerry, order = 2)
```

We can save the original weights:

```{r}
queen_w <- queen_weights(guerry)
```

Then take a look at the class:
```{r}
class(queen_w)
```

```{block type="learncheck"}
**Question**
```
Try out the following functions:

- `queen_w$is_symmetric`
- `min_nbrs`
- `max_nbrs`
- `GetNeighbors(0)`

Find more listed here: [Attributes of `Weight` object](https://rgeoda.github.io/rgeoda-book/spatial-weights.html#attributes-of-weight-object)
```{block type="learncheck"}
```

Many more types of weights can be found [here](https://rgeoda.github.io/rgeoda-book/spatial-weights.html#attributes-of-weight-object).

### Detour: saving your weights

To save contiguity-based weights, use the `SaveToFile()` function in rgeoda.

The four arguments you need for `SaveToFile()` are the:

- path/name of the file you're going to create (ends in .gal for contiguity-based weights, .gwt for distance-based weights)
- layer name (aka file name without suffix, ie XXX.shp )
- id column name, a unique value for each observation
- id column itself

```{r eval=FALSE}
id_names <- guerry$GetIntegerCol("CODE_DE")
queen_w$SaveToFile("guerry_queen.gal", "Guerry", "CODE_DE", id_names)
```

## Calculating the Local Moran statistic

You can then calculate the Local Moran statistic once you have your weights:

```{r}
crm_prp <- guerry_sf$Crm_prp

lisa <- local_moran(queen_w, crm_prp)
```

And make a map!

```{r}
lisa_colors <- lisa$GetColors() 
lisa_labels <- lisa$GetLabels()
lisa_clusters <- lisa$GetClusterIndicators()

plot(st_geometry(guerry_sf), 
     col=sapply(lisa_clusters, function(x){return(lisa_colors[[x+1]])}), 
     border = "#333333", lwd=0.2)
title(main = "Local Moran Map of Crm_prp")
legend('bottomleft', legend = lisa_labels, fill = lisa_colors, border = "#eeeeee")
```


